### 2019-04-15 paper/paper.tex
updated bibs, more ecosystem content

In addition to the number of external dependencies per package, 
we investigate how packages support dependants. The original paper 
discovered that the majority of in degrees are concentrated among a core 
minority of packages, with the majority of packages having no dependants,
a discovery that has also been observed in other software ecosystems such
as Linux and MySQL \cite{Myers:2003}. In our updated data, we discover 
that not only have these aspects held true, but the concentration of 
dependencies has actually increased. As of April 1\textsuperscript{st} 2019, [NUMBER]\% of packages
had zero dependants, up from [NUMBER]\% in 2015. Meanwhile, only \%[NUMBER] of packages
had 6 or more dependants, down from \%[NUMBER] in September 2015. It is again interesting to note
that the current trend of increasing concentration appears to have begun shortly after the end of window in the original
paper \cite{Wittern:2016}.

@@

\begin{figure}
  % TODO NOTE THAT THIS HAS LOWER NUMBERS THAN THE ORIGINAL PAPER DUE TO MONTHLY PAGERANKS
  \begin{tabular}{c|c|c|c}
    Year & Top 10 & Top 100 & Top 250 \\
    \hline
    2010 & 13 & 103 & 253\\
    2011 & 25 & 297 & 701\\
    2012 & 6 & 46 & 136\\
    2013 & 3 & 38 & 114\\
    2014 & 1 & 40 & 96\\
    2015 & 4 & 38 & 86\\
    2016 & 4 & 29 & 62\\
    2017 & 0 & 12 & 48\\
    2018 & 0 & 15 & 31\\
    2019 & 0 & 4 & 12\\
  \end{tabular}
  \caption{Number of packages entering top \code{npm} ranks for the first time}
  \label{numEnteringTop}
\end{figure}

\section{Package Popularity}

As in the original paper we analyzed popularity of \code{npm} packages using three major measures
\begin{enumerate}
  \item The \textbf{\code{npm} rank} is calculated using the PageRank\cite{brin1998anatomy} of an \code{npm} package within a dependency graph that we built using the
    package metadata as described in Section \ref{PackageMetadata}. PageRank is calculated for every package using a damping factor of 0.85 
    and stopping iterations once the cumulative change in values of vertices went below $10^{-6}$. We then order all packages by their PageRank to determine the
    \code{npm} rank, an relative ranking of package popularity with 1 being the most popular.
  \item The \textbf{download rank} is based on the the gross volume of package downloads from the \code{npm}
    registry, then ordered relatively with 1 being the most downloaded.
  \item \textbf{\code{GitHub} rank} is a measure of the popularity of package use in public \code{GitHub} projects.
    This rank is created from a list of \code{npm} dependency occurrences found in \code{package.json} files, with rank 1
    meaning that the package was the most common dependency among analyzed \code{GitHub} repositories.  
\end{enumerate}

@@

\subsection{Popularity Over Time}
[NEEDS MORE DATA]

@@

>>>>

@@

In addition to the number of external dependencies per package, 
we investigate how packages support dependants. The original paper 
discovered that the majority of in degrees are concentrated among a core 
minority of packages, with the majority of packages having no dependants,
a discovery that has also been observed in other software ecosystems such
as Linux and MySQL \cite{Myers:2003}. In our updated data, we discover 
that not only have these aspects held true, but the concentration of 
dependencies has actually increased. As of April 1\textsuperscript{st} 2019, [NUMBER]\% of packages
had zero dependants, up from [NUMBER]\% in 2015. Meanwhile, only \%[NUMBER] of packages
had 6 or more dependants, down from \%[NUMBER] in September 2015. It is again interesting to note
that the current trend of increasing concentration appears to have begun shortly after the end of window in the original
paper.

In section \ref{popularity}, we revisit the popularity measures
investigated in the original paper \cite{Wittern:2016} to discover
changes in the relationships between package rankings
both inside and outside of the \code{npm} ecosystem,
gaining insight into the dynamics of \code{npm} package popularity.

\textbf{Takeaways.} We find that while the \code{npm} ecosystem
continues to grow, the rate of growth is no longer superlinear, 
a possible sign of the ecosystem beginning to reach some form of maturity.
In the relationships between packages, we find that though the average number
of dependencies continues to grow, there has been a recent decline in the
rate of this growth. Accordingly, we have found that the number of packages with
zero dependencies has increased from 28.7\% in September 2015 to [DATA REQUIRED]\%
as of April 2019. The percentage of packages that are depended on has
also declined, albeit still exhibiting a power law relationship as observed
in prior software ecosystem research \cite{Wittern:2016, Myers:2003, Louridas:2008}.

\section{Package Popularity}\label{popularity}

As per the original paper we analyzed popularity of \code{npm} packages using three major measures \cite{Wittern:2016}:
\begin{enumerate}
  \item The \textbf{\code{npm} rank} is calculated using the PageRank\cite{brin1998anatomy} of an \code{npm} package within a dependency graph that we built using the
    package metadata as described in Section \ref{PackageMetadata}. PageRank is calculated for every package using a damping factor of 0.85 
    and stopping iterations once the cumulative change in values of vertices went below $10^{-6}$. We then order all packages by their PageRank to determine the
    \code{npm} rank, an relative ranking of package popularity with 1 being the most popular.
  \item The \textbf{download rank} of a package is based on the the gross volume of package downloads from the \code{npm}
    registry, then ordered relatively with 1 being the most downloaded.
  \item \textbf{\code{GitHub} rank} is a measure of the popularity of a package among public \code{GitHub} projects.
    This rank is created from a list of \code{npm} dependency occurrences found in \code{package.json} files, with rank 1
    meaning that the package was the most common dependency among analyzed \code{GitHub} repositories.  
\end{enumerate}

@@

\subsection{Popularity Over Time}

\begin{table}
  \centering
  % TODO NOTE THAT THIS HAS LOWER NUMBERS THAN THE ORIGINAL PAPER DUE TO MONTHLY PAGERANKS
  \begin{tabular}{c|c|c|c}
    Year & Top 10 & Top 100 & Top 250 \\
    \hline
    2010 & 13 & 103 & 253\\
    2011 & 25 & 297 & 701\\
    2012 & 6 & 46 & 136\\
    2013 & 3 & 38 & 114\\
    2014 & 1 & 40 & 96\\
    2015 & 4 & 38 & 86\\
    2016 & 4 & 29 & 62\\
    2017 & 0 & 12 & 48\\
    2018 & 0 & 15 & 31\\
    2019 & 0 & 4 & 12\\
  \end{tabular}
  \caption{Number of packages entering top \code{npm} ranks for the first time}
  \label{numEnteringTop}
\end{table}

[NEEDS MORE DATA]

@@

  \bibitem{Louridas:2008}
  Panagiotis Louridas, Diomidis Spinellis, and Vasileios Vlachos.
  \newblock Power laws in software.
  \newblock {\em ACM Trans. Softw. Eng. Methodol.}, 18(1):2:1--2:26, October
    2008.
